---
title: "Data Cleaning Assignment"
author: "Eureka Project Team"
date: "10/23/2019"
output:
  html_document:
    theme: cosmo
    toc: yes
    code_folding: hide
  github_document:
    toc: yes
---

```{r setup, echo=TRUE, message=FALSE, warning=FALSE} 
knitr::opts_chunk$set(cache=TRUE)
knitr::opts_knit$set(root.dir = "~/Documents/ISQA8086/R_scripts/") 
```
# Description of Xerces Data Set

The project team was provided the data in Excel format. The Excel spreadsheet was composed of six sheets, Site, Visit, Veg (Vegetation), Macro, Seed and Species Attributes. Below you will find descriptions of each of the Excel sheets. We did not have any intellectual property constraints or licensing issues associated with the data.  For this Rmarkdown script to work you will need to download the site_data.csv and landscape.csv.

## Data Reference

Powers, R 2019, *Plant establishment in Conservation Reserve Program's pollinator habitats across Nebrasksa*, data file, NRCS-Xerces Society, *Microsoft Excel file*

## Site Data Description

The Site data is composed of 17 rows and 10 columns. It contains a field id, a region, the number of acres associated with the field, the county, the date the site was planted, the planting method, the number of species planted, the Pure Live Seed pounds per acre, the previous vegetation cover and the Vendor for the seed planted. Only two cells were left N/A. 

The Columns for the Site Data: 

<table border=2> 
<th> Field ID </th>
<th> Region </th>
<th> Field acres </th>
<th> County </th>
<th> Planted Date </th>
<th> Planting method </th>
<th> # of species planted </th>
<th> Pure Live Seed lbs/acre </th>
<th> Previous cover </th>
<th> Seed Vendor </td>
</table>
<hr/>

### Remediation of the Site Data

On a whole the site data was fine. There were two missing cells for Number of Species and Pure Live Seed pounds per acre for the Prairie Plains Seed Vendor. Otherwise, the data set was complete. From the planted date, we will look at deriving a new field called PlantingSeason. The PlantingSeason allows for a friendlier translation of the date. However, the Site data alone is not enough information to draw many conclusions. As a result, it will be necessary to merge this data with the visit data to increase its value. The planting method will be removed since the method is drill for all sites.

## Visit Data Description  

The Visit data is composed of 56 rows and 7 columns. It contains a field ID, a visit ID, Transect, the visit date, the observer, the surrounding Landscape and the Notes about the visit. Much of the data for the surrounding landscape is duplicated for each transect and each visit. The notes cells have the missing data for this sheet. The notes also contain an indication if the site failed. Another indication of a failed site is a fall visit will not occur. The biggest issue with this data are the Surrounding Landscape and the Notes columns. These columns are free text fields and difficult to parse. The Surrounding Landscape data has the potential to add value though. As a result, this data was parsed manually with a description of this new dataset below.  

The Columns for the Visit Data:  

<table border=2> 
<th> Visit ID </th>
<th> Field ID </th>
<th> Transect </th>
<th> Date </th>
<th> Observer </th>
<th> Surrounding Landscape </th>
<th> Notes </td>
</table>
<hr/>

The columns for the parsed Surrounding Landscape Data:

<table border=2> 
<th> Field ID </th>
<th> LandSTree </th>
<th> LandSRoad </th>
<th> LandSStructuresr </th>
<th> LandSSCrop </th>
<th> LandSGrass </th>
<th> LandSWaterArea </th>
<th> LandSOther </td>
</table>
<hr/>  

### Remediation of the Visit Data

The visit data, on a whole, was fine. Most of the data was present. Only the Notes columns had missing cells and that is to be expected for a free form field. The Notes data for the most part will be ignored since it is more of a qualitative data. Although, the Notes column does tell us if the site was a failure. We will take that information and manually create a new column called established and populate the words success or failure. The Surrounding Landscape data was free form data as well. At first site we might have categorized this data as qualitative as well but the client specifically called out this data as variable in the analysis. The current form the data was unusable though. Using Excel, this data was parsed into 7 columns. The data was collapsed to the Field ID level. This collapsing was done since the data did not vary by transect, observer or visit date. The observer was Ms. Rae Powers for every visit except two where she was joined by another individual.

#### Steps  

1. Created copy of Visit data Excel Sheet and named it visit_data_updated  

2. Added new columns to Excel sheet (LandSTree,LandSRoad,LandSStructures,LandSSCrop,LandSGrass,LandSWaterArea,LandSOther, Established)  

3. Added data to each of the new columns based on the contents in the Surrounding Landscape column  

4. Consolidated all visits into one row for each Field ID and removed extra rows.  

5. Removed Visit ID, Observer, Visit Date, Transect, Surrounding LandScape and Notes columns.  

6. Merged visit_data_updated sheet with site data sheet based on Field ID  


The New consolidated dataset, ExpandedSiteData

<table border=2> 
<th> FieldID </th>
<th> Region </th>
<th> NumOfAcres </th>
<th> County </th>
<th> PlantedDate </th>
<th> PlantingSeason </th>
<th> SpeciesNum </th>
<th> TotalPLSAcre </th>
<th> PreCover </th>
<th> SeedVendor </th>
<th> LandSTree </th>
<th> LandSRoad </th>
<th> LandSStructuresr </th>
<th> LandSSCrop </th>
<th> LandSGrass </th>
<th> LandSWaterArea </th>
<th> LandSOther </th>
<th> Established </td>
</table>
<hr/>  

### Load Site and Landscape Data  

```{r loadData, echo=FALSE}
site <- read.csv("site_data.csv")
landS <- read.csv("landscape.csv")
sitelandS <- merge(x=site, y=landS, by="FieldID", all=TRUE)
print(sitelandS)
```
## Veg Data Description

The Veg data is composed of 891 rows and 22 columns. The data set contains a Visit ID, a species and 20 quadrat columns divided by floral cover and non-floral cover. The quadrat contains a number from 1 to 6 that translates to the amount of coverage for that species within the quadrat. Much of this data is blank but not missing which is to be expected.  

The Columns for Veg Data:  

<table border=2> 
<th> Visit ID </th>
<th> Species </th>
<th> Q1 </th>
<th> Q1FC </th>
<th> Q2 </th>
<th> Q2FC </th>
<th> Q3 </th>
<th> Q3FC </th>
<th> Q4 </th>
<th> Q4FC </th>
<th> Q5 </th>
<th> Q5FC </th>
<th> Q6 </th>
<th> Q6FC </th>
<th> Q7 </th>
<th> Q7FC </th>
<th> Q8 </th>
<th> Q8FC </th>
<th> Q9 </th>
<th> Q9FC </th>
<th> Q10 </th>
<th> Q10FC </td>
</table>
<hr/>  

### Remediation of Veg Data  

The Veg data requires some additional data from the visit dataset and the site data set. Below are the steps necessary to merge that data.

1. Export each tab of excel spreadsheet into individual csv
      a. File > Export > Change File Type > CSV > Save As  
      b. "Site data" saved as "site data.csv"  
      c. "Visit data" saved as "visit data.csv"  
      d. "Veg data" saved as "veg data.csv"  
      e. "Species attributes" saved as "species attributes.csv"         
2. Import each csv into RStudio  

3. Combine all the data  

4. Trim the unneeded data columns  

## Macro Data Description  

The Macro data is composed of 650 rows and 5 columns. The data set contains a Visit ID, 3 columns that categorizes the species cover in the macroplots, <1%, 1-5% and > 5%, and Notes. The Notes are free form text. Much of this data is blank but not missing which is to be expected.  

The columns for Macro Data:  

<table border=2> 
<th> Visit ID </th>
<th> <1% </th>
<th> 1-5% </th>
<th> 5%> </th>
<th> Notes </td>
</table>
<hr/>

## Seed Data Description  

The Seed data is composed of 543 rows and 4 columns. The data set contains a Field ID, Species, Pure Live Seed lbs per acre and Origin. This data set contains no missing data. The sum of the Pure Live Seed pound per acre by Field ID data matches the value found for that Field ID in the Site data sheet.  

The columns for Seed Data:  

<table border=2> 
<th> Field ID </th>
<th> Species </th>
<th> Pure Live Seed lbs/acre </th>
<th> Origin </td>
</table>
<hr/>
 

## Species Attributes Description  

The Species attributes data is composed of 182 rows and 3 columns. The data set contains Species, Coefficient of Conservatism and common Name. A significant portion of the Coefficient of Conservatism is marked with an asterisk. These species will not be included in any coefficient of conservation averages per the client.  

The Columns for Species Attributes Data:  

<table border=2> 
<th> Species </th>
<th> Coefficient of Conservatism </th>
<th> Common Name </td>
</table>
<hr/>

## Data Dictionary
<table border=2> 
<th>Attribute Name</th>

<th>Attribute Description</td>
<tr>
<td>FieldID</td>
<td>The name assigned to the site</td>
<tr> 
<td>Region</td>
<td>The Region of the state, like NorthEast(NE)</td>
<tr> 
<td>NumOfAcres</td> 
<td>The number of Acres associated with the site</td> 
<tr>
<td>County </td> 
<td>The county of the site</td> 
<tr>
<td>PlantedDate</td> 
<td>The date the species were planted</td> 
<tr>
<td>PlantingSeason</td> 
<td>The season when the species were planted </td> 
<tr>
<td>Established</td> 
<td>The success or failure of the site </td> 
<tr>
<td>VisitNum	</td> 
<td>The Number of Visits </td> 
<tr>
<td>SpeciesNum	</td> <td>The number of Species planted </td> <tr>
<td>TotalPLSAcre	</td> <td>The total amount of Pure Live Seed by pound by acre planted at the site </td> <tr>
<td>PreCover	</td> <td>The species that was the previous cover </td> <tr>
<td>SeedVendor	</td> <td>The name of the vendor that provided the seeds </td> <tr>
<td>Transect </td> <td>The areas in which the site is divided. In most cases there are two transect but in one case there are three. </td> <tr>
<td>LandSTree	</td> <td>Contains the trees in the surrounding landscape </td> <tr>
<td>LandSRoad	</td> <td>Indicates if there is a road in the surrounding landscape </td> <tr>
<td>LandSStructures	</td> <td>  Indicates if there are housing in the surrounding landscape </td> <tr>
<td>LandSCrop	 </td> <td> Indicates what types of crops are in the surrounding landscape </td> <tr>
<td>LandSGrass	 </td> <td> Indicates what type of grass is in the surrounding landscape </td> <tr>
<td>LandSWaterArea	</td> <td>Indicates that some source of water exists near the site </td> <tr>
<td>LandSOther	</td> <td>Indicates an attribute of the surrounding landscape that did not fall within the other 6 categories </td> <tr>
<td>VisitID   </td> <td> Indicator to distinguish visits; starts with SP for Spring  or F for Fall and contains 1,2, or 3 for the transect </td> <tr>
<td>VisitDate  </td> <td> Indicates date of the visits </td> <tr>
<td>VisitObserver  </td> <td>Indicates who made the observations during the visits </td> <tr>
<td>SurrLandScape </td> <td>Notation like text about the surrounding Landscape of the site; this data was broken up for more usability </td> <tr>
<td>VisitNotes        </td> <td> Free form text about the visits </td> <tr>
<td>Species          </td> <td> The species that were planted at the site or were observed at the site </td> <tr>
<td>SpeciesOrigin   </td> <td>  The origin of the Species planted </td> <tr>
<td>SpeciesPLS      <td>   The pure live seed pounds per acre that was planted for a species </td> <tr>
<td>SpeciesCommName  </td> <td>   The Common Name of a Species </td> <tr>
<td>CoEOfConsvm		</td> <td>	The Coefficient of Conservatism which is a range between 0 and 10. The higher the score the lower tolerance to environmental degradation </td> 
<tr>
</table>
<hr/>  

### Contributionship Statement  
Kristi worked on the Remediation for the Site and Visit Data and the Data Dictionary  
Erik worked on the Remediation for the Veg data  
Anitha was the Proofreader  

### Proofreader Statement
I Anitha Saminathan have reviewed and proofread the following work and agree that it is ready for submission.
